
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing required libraries:

```{r 1}

library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
#
# color-blind palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

lvB <- read.csv('/Users/aneeshnaik/Downloads/lvBetween_25oct18.csv')

head(lvB)
```

## Recoding conditions:

Before we begin our analysis we must recode the conditions into factors of interest:

```{r 2}
lvB$amb <- ifelse(lvB$newCond == "a" |
                   lvB$newCond == "d", "amb",
                 ifelse(lvB$newCond == "b" |
                   lvB$newCond == "e", "low", "high"))

lvB$firstGen <- ifelse(lvB$newCond == "a" |
                   lvB$newCond == "b" |
                   lvB$newCond == "c", "f", "m")
lvB$secondGen <- ifelse(lvB$newCond == "a" |
                   lvB$newCond == "e" |
                   lvB$newCond == "f", "f", "m")

str(lvB)

lvRel <- lvB[lvB$pos >= 4 & lvB$pos <= 8,]
str(lvRel)

```

## Plots:

Now in order to get a sense of the distribution of the data, we make the following plots:

```{r 3}

# reading times across all conditions:

lvP0 <- ggplot(lvRel,
                aes(pos, RT))

lvP1 <- lvP0 + stat_summary(fun.y = mean, geom = "line", size = 1) 

lvP1

# separately for the three ambiguity values:

lvP0 <- ggplot(lvRel,
                aes(pos, RT,
                    colour = factor(amb)))

lvP1 <- lvP0 + stat_summary(fun.y = mean, geom = "line", size = 1) 

lvP1

# separately for the gender of the first noun as a function of the gender of the second noun:

lvP0 <- ggplot(lvRel,
                aes(pos, RT,
                    colour = factor(firstGen)))

lvP1 <- lvP0 + stat_summary(fun.y = mean, geom = "line", size = 1) +
  facet_wrap(~ secondGen)

lvP1

# and finally putting it all into a single plot, by ambiguity and gender of the first noun:

lvP0 <- ggplot(lvRel,
                aes(pos, RT,
                    colour = factor(amb),
                    linetype = firstGen))

lvP1 <- lvP0 + stat_summary(fun.y = mean, geom = "line", size = 1)

lvP1
```

## ANOVA:

We are now ready to perform the ANOVA that tests the influence of the factor values on reading times at a given position. 

```{r 4}
# Calculating means:

subjectMeans <- lvRel %>% 
  filter(pos == 6) %>%
  group_by(newSubj, amb, firstGen) %>%
  summarise(meanRT = mean(RT)) %>%
  data.frame()

colnames(subjectMeans) <- c("subj",
                            "amb",
                            "firstGen",
                            "meanRT") 


grandMean <- mean(subjectMeans$meanRT)
g_means <- tapply(subjectMeans$meanRT, subjectMeans$firstGen, mean)
pos_means <- tapply(subjectMeans$meanRT,subjectMeans$amb, mean)
g_pos_means <- tapply(subjectMeans$meanRT, list(subjectMeans$firstGen,subjectMeans$amb), mean)

grandMean;g_means;pos_means;g_pos_means

# Calculating sums of squares:

sos_total <- sum((subjectMeans$meanRT - grandMean)^2)
sos_total

sos_gen <- sum((g_means - grandMean)^2)
sos_gen <- sos_gen*24*3

sos_amb <- sum((pos_means-grandMean)^2)
sos_amb <- sos_amb*24*2

acc <- 0
for(i in 1:2){
  for(j in 1:3){
    acc = acc + (g_pos_means[i,j] - g_means[i] - pos_means[j] + grandMean)^2
  }
}
sos_g_amb <- 24*sum(acc)
sos_g_amb

sos_err <- sos_total-sos_gen-sos_amb-sos_g_amb
sos_err

# Calculating mean square errors:

ms_total <- sos_total/(144-1)
ms_gen <- sos_gen/(2-1)
ms_amb <- sos_amb/(3-1)
ms_gen_amb <- sos_g_amb/((2-1)*(3-1))
ms_err <- sos_err/(144-2*3)

# Calculating F stats:

f_gen <- ms_gen / ms_err
f_amb <- ms_amb / ms_err
f_g_amb <- ms_gen_amb/ ms_err

# Comparing results to output of aov():

f_gen;f_amb;f_g_amb

ms_total;ms_gen;ms_amb;ms_gen_amb;ms_err

sos_total;sos_gen;sos_amb;sos_g_amb;sos_err

summary(aov(meanRT ~ firstGen + amb + firstGen:amb,
            data = subjectMeans))

```

We see that our handmade ANOVA produces the same results as the aov() function. From the p values it seems like the gender of the first noun and the attachment position have some effect on the reading times as the p values are < 0.05. The interaction factor does not have a significant effect on reading times. 


## Tukey's HSD:
```{r 5}

anova <- aov(meanRT ~ firstGen + amb + firstGen:amb,
            data = subjectMeans)

TukeyHSD(anova)

plot(TukeyHSD(anova),las = 1,cex.axis = 0.75)
```

We see that the largest and most significant pair-wise comparisons (p val < 0.05) are between 'm' and 'f' for the gender of the first noun, between the low and ambiguous attachment positions and the interaction of these two which would be the 'm:low-f:amb' comparison. 
